A Framework for Classification of Social Machines

### 2013-02-15 latex/sig-alternate.tex
Added "Open Access" as a SM, and updated figure to show it

Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behaviours at various levels of granularity.  We propose looking at nested machines with ``The Web'' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahid platforms) that spawn more specific social machines. A resulting polyarchy is shown in Figure \ref{polyarchy}, which classifies levels as Infrastructure, Frameworks, Services and Causes/Groups.

@@

The sub constructs are illustrated in Figure \ref{motivation}, and we relate them to Maslow's hierarchy of needs~\cite{maslow}, (e.g., self-actualisation, security, respect of others), in various ways.

@@

Finally, we identified several constructs that are relevant for all social machines. First, there are construct that relate to ``analytics'' that can be measured at a point in time, and may change over time. We have documented these constructs in Table \ref{table:constructs}. In the left column we list construct regarding the design affordances of the social machine. In the middle column we have a mixture of the result of how the design affordances actually were realised by the social computation, such as how users co-opted / incorporated / interpreted the constraints. Finally, in the right column we list analytical measures of social machines which are subject to change over time.

@@

>>>>

Looking at the set of social machines in a polyarchy leads to a broad/specific relationship emerging that lets us talk about behaviours at various levels of granularity.  We propose looking at nested machines with ``The Web'' as one of several potential roots, with the next level down consisting of sub-platforms (such as Facebook, Twitter, MediaWiki, Ushahid platforms) that spawn more specific social machines. A resulting polyarchy is shown in Figure~\ref{polyarchy}, which classifies levels as Infrastructure, Frameworks, Services and Causes/Groups.

@@

The sub constructs are illustrated in Figure~\ref{motivation}, and we relate them to Maslow's hierarchy of needs~\cite{maslow}, (e.g., self-actualisation, security, respect of others), in various ways.

@@

Finally, we identified several constructs that are relevant for all social machines. First, there are construct that relate to ``analytics'' that can be measured at a point in time, and may change over time. We have documented these constructs in Table~\ref{table:constructs}. In the left column we list construct regarding the design affordances of the social machine. In the middle column we have a mixture of the result of how the design affordances actually were realised by the social computation, such as how users co-opted / incorporated / interpreted the constraints. Finally, in the right column we list analytical measures of social machines which are subject to change over time.

@@

\subsection{Causes as Social Machines}

One originally unforeseen classification is that of the causes that utilise Social Machines.
Through a particularly popular cause, large numbers of participants can be mobilised, for
either a short period of time, or for long-term engagement, using multiple social machines
in order to further their cause. Examples of causes can be seen in Figure~\ref{polyarchy},
at the lowest level of the polyarchy. One such example is the cause of ``Open Access'' to
academic publications~\cite{harnad2001self}. This cause mobilises academic authors to
self-archive their own papers online so they are available for free, in addition to hosted
versions behind publishers pay-walls. In order to achieve this aim, the social machine of
``open access'' typically uses the social machines of institutional repositories provided
by the authors' associated institutions. As with other social machines, these social
machines typically use off-the-shelf implementations (in this case the most popular are
EPrints~\cite{eprints} and DSpace~\cite{dspace}), which run as web sites on the world wide
web. This cause is mature, growing year-on-year and has global coverage.

[more about this?]
